\begin{appendices}

\chapter{Self-Appraisal}
\label{appendix:selfappraisal}

\section{Project process reflection}

The phased gate structure was the single most effective process decision in the project. Each phase's result determined the next phase's design rather than following a pre-specified plan. Phase~2's finding that rainbow-lite was the only collapse-resistant algorithm in single-agent evaluation meant that Phase~3 could attribute multi-agent collapse to MARL non-stationarity rather than to algorithm fragility. Without this gate, the Phase~3 collapse results would have been ambiguous. The same principle applied at each subsequent gate. Phase~4's zero-sum failure motivated Phase~5's game structure change. Phase~5's success motivated Phase~6's scaling test.

The config-driven fairness control was also effective. Because all experimental parameters were externalised to \texttt{config.json} from the start, every comparison was provably identical in all conditions other than the variable under test. An external reviewer can reconstruct any trial from the config snapshot embedded in the metric output file. This discipline was not costly to implement but prevented a class of confounds that would have been difficult to detect and correct retrospectively.

The decision to continue beyond Phase~3 rather than stopping at the first successful result produced the most original contributions. The 83 per cent cooperative improvement in Phase~5 would not have been found if the project had ended at the Phase~3 competitive equilibrium results. The scaling failure in Phase~6, the boundary characterisation in Phase~7, the curriculum decomposition in Phase~8, and the difference rewards analysis in Phase~10 all followed from the decision to keep asking the next question.

The algorithm confound in Phase~10 was the main area that did not work well. The \texttt{config.json} file still contained \texttt{"algo": "vanilla"} when rainbow-lite was intended for Phase~10A. The first batch of trials therefore tested the wrong algorithm. The error was caught by reading the \texttt{"algorithm"} field embedded in the output metric JSON rather than by reviewing the config file before running. The lesson is that config inspection at the start of every experiment batch is a necessary check, not an optional one. Embedding config snapshots in telemetry output is an effective safeguard, but it should not be the only one.

Writing pace was a second area for improvement. Chapters~1 to~3 were drafted while Phases~1 to~3 were active, which made them detailed and timely. Phases~4 to~10 were completed before the corresponding chapters were written, which made the retrospective write-up harder. Writing contemporaneously with experimentation is a better practice.

\section{Personal reflection and lessons learned}

The most significant conceptual shift during the project was the recognition that negative results are contributions. Phase~10's failure to produce cooperative behaviour via difference rewards was frustrating initially. The mathematical analysis of the gradient inversion transformed it into a finding. The formula failed. The reason it failed can be stated precisely. This precision is what makes a negative result publishable rather than merely disappointing.

The N of four scaling cliff was not predicted. The project could have treated Phase~6's failure as a dead end and stopped. Designing Phase~7 to characterise the boundary precisely was the decision that produced a genuinely new empirical finding. The habit of asking "where exactly does this transition occur?" rather than "does this work?" produced more informative results throughout the project.

Reproducibility standards imposed discipline that paid dividends later. Because all prior phases used fixed seeds and matched budgets, the Phase~7 and Phase~8 ablations could attribute results to game structure and training schedule rather than to evaluation noise. The upfront cost of implementing this discipline was small. The downstream benefit of being able to make confident comparative claims was large.

\section{Legal, social, ethical and professional issues}

\subsection{Ethical issues}

No human participants were involved at any stage of this project. No consent forms were required. The LLM baseline in Phase~9 uses the Anthropic Claude Haiku API. No personal data was transmitted to the API at any point. Race state information is synthetic and contains no identifying information.

\subsection{Legal issues}

All external libraries used in this project are open-source. FastF1 is released under the MIT licence. PyTorch is released under a BSD-style licence. NumPy and Pandas are released under BSD licences. The Anthropic Claude Haiku API was used under Anthropic's standard commercial terms of service. The FastF1 telemetry data is drawn from publicly available Formula~1 race data intended for non-commercial research use, consistent with the library's intended purpose and licence terms.

\subsection{Social issues}

This project does not interact with human participants and does not collect or process personal data. The research question concerns the behaviour of autonomous software agents in a simulated environment. There are no direct social impacts from the project itself.

The broader research area has social relevance. Autonomous agents with configurable incentive structures are deployed in domains beyond motorsport simulation, including financial trading, logistics optimisation, and competitive resource allocation. The finding that incentive formula design shapes emergent agent behaviour as strongly as algorithm choice is relevant to anyone designing the reward contracts for multi-agent systems in these domains. The project provides controlled evidence that poorly designed cooperative incentives can produce competitive or even harmful agent behaviour, which is a meaningful contribution to the broader conversation about AI alignment in multi-agent settings.

\subsection{Professional issues}

This project follows the reproducibility standards identified as necessary for trustworthy reinforcement learning comparisons by \citet{henderson2018deep} and \citet{agarwal2021statistical}. Matched compute budgets, fixed seeds, confidence interval reporting, and behavioural diagnostics that go beyond outcome-only metrics are treated as professional obligations rather than optional additions. The git repository provides a full audit trail of experimental code, configuration changes, and metric outputs. All claims in the main report are traceable to specific metric files or mathematical derivations.

\chapter{External Materials}
\label{appendix:external}

All materials not developed by the student are listed below.

\begin{table}[htbp]
\centering
\small
\caption{External materials used in this project}
\label{tab:external-materials}
\begin{tabular}{|p{3.5cm}|p{2.5cm}|p{5cm}|p{2.5cm}|}
\hline
\textbf{Material} & \textbf{Version} & \textbf{Purpose} & \textbf{Licence} \\
\hline
FastF1 & v3.x & 2023 Belgian Grand Prix telemetry for overtaking zone difficulty calibration & MIT \\
\hline
PyTorch & v2.x & Neural network implementation for all DQN-family variants & BSD \\
\hline
NumPy & v1.x & Numerical computation in simulator and metrics scripts & BSD \\
\hline
Pandas & v2.x & Metrics aggregation and results table generation & BSD \\
\hline
Anthropic Claude Haiku & API (2024) & Zero-shot LLM agent for Phase~9 semantic baseline & Commercial API terms of service \\
\hline
\end{tabular}
\end{table}

All academic papers cited in this report were accessed through publicly available sources or the University of Leeds library. No software developed for this project relies on proprietary algorithms or datasets. The project codebase is available to the supervisor and assessor via the git repository linked in the project submission.

The related work comparison table from the original report is reproduced in Table~\ref{tab:related-work-comparison} below for reference.

\begin{table}[htbp]
\centering
\scriptsize
\caption{Comparison of related work and relevance to this dissertation}
\label{tab:related-work-comparison}
\resizebox{\textwidth}{!}{
\begin{tabular}{|p{2.7cm}|p{2.8cm}|p{2.7cm}|p{2.5cm}|p{2.5cm}|p{2.7cm}|p{2.8cm}|}
\hline
\textbf{Study} & \textbf{Domain and simulator fidelity} & \textbf{Decision method} & \textbf{Uncertainty treatment} & \textbf{Evaluation rigour} & \textbf{Limitations} & \textbf{Direct relevance to thesis} \\
\hline
Heilmeier et al. (2018) & Circuit motorsport with lap-wise race simulation and tyre and fuel effects & Strategy simulation and what-if analysis & Includes overtaking and degradation effects in model dynamics & Practical validation on historical race setting & Not focused on RL benchmark fairness across seeds & Motivates staged simulator design with tyre extension \\
\hline
Heilmeier et al. (2020) & Circuit motorsport strategy support with ANN components & Surrogate modelling and optimisation support & Captures race variability through modelled scenarios & Strong engineering framing and deployability discussion & Limited focus on RL reproducibility protocol & Supports need for interpretable strategy tools \\
\hline
Piccinotti et al. (2021) & Formula 1 strategy identification in planning setting & Online planning methods & Dynamic race context in planning loop & Demonstrates value of planning for strategy choice & Limited DQN-family comparative analysis & Clarifies boundary between planning and RL benchmarking goals \\
\hline
Liu and Fotouhi (2020) & Formula E race strategy with energy constraints & ANN plus MCTS & Multi-stage uncertainty in energy and race conditions & Demonstrates decision quality in time-critical domain & Not designed as controlled RL reproducibility benchmark & Supports tactical decision framing under uncertainty \\
\hline
Liu et al. (2024) & Formula E multi-car strategy learning context & Reinforcement learning based strategy development & Competitive race uncertainty in multi-car setting & Strong domain result emphasis & Less emphasis on seed-controlled statistical protocol & Provides adjacent motorsport RL evidence base \\
\hline
Mnih et al. (2015), van Hasselt et al. (2016), Wang et al. (2016), Hessel et al. (2018) & General RL benchmarks & DQN, Double DQN, Dueling DQN, Rainbow & Algorithmic robustness mechanisms in value learning & Benchmark-oriented algorithm studies & Not motorsport-specific and limited tactical interpretability & Defines core algorithm comparison set for this dissertation \\
\hline
Henderson et al. (2018), Agarwal et al. (2021) & Cross-domain deep RL evaluation & Reproducibility and statistical evaluation methodology & Focus on stochastic outcome variation and seed effects & Strong critique of weak reporting practice & Do not provide motorsport-specific protocol & Directly motivates fixed-seed and interval-based evaluation design \\
\hline
Sankary and Ostfeld (2018) & Robust optimisation in water systems & Evolutionary optimisation with stochastic scenario selection & Explicit stochastic scenario resampling & Robustness-focused evaluation method & Different optimisation paradigm from RL & Supports uncertainty-aware evaluation discipline \\
\hline
This dissertation & F1-inspired competitive simulator with staged complexity & Reproducible DQN-family benchmark with behavioural diagnostics & Controlled stochasticity pathway with tyre-related uncertainty stage & Matched budget and seed-controlled reporting with confidence intervals & Simplified simulator abstraction and staged feature scope & Targets interpretable and statistically grounded method comparison \\
\hline
\end{tabular}
}
\end{table}


\chapter{Phase 0 and Phase 1 Experimental Validation}
\label{appendix:phase01}

This appendix documents the system integrity checks carried out in Phase~0 and the smoke evaluation results collected in Phase~1. Both phases are prerequisites to the Phase~2 DQN-family benchmark. Phase~0 verified that the simulator, reward accounting, stochasticity, and telemetry were correct before any learning claims were made. Phase~1 confirmed a learning signal at low compute cost and established the stochasticity sensitivity profile of the vanilla DQN baseline.

% ---------------------------------------------------------------
\section{Phase 0: System Integrity}
% ---------------------------------------------------------------

\subsection{Objective and scope}

Phase~0 contained twelve white-box unit and integration tests grouped into three categories. Group~A checked simulator correctness. Group~B validated the stochastic probability calculations statistically. Group~C verified the telemetry contract. All tests were implemented in \texttt{tests/test\_phase0\_integrity.py} and executed under a fixed random seed with the Spa circuit at low complexity and stochasticity level s0.

\subsection{Test inventory}

\begin{table}[htbp]
\centering
\scriptsize
\caption{Phase 0 integrity test inventory and outcomes}
\label{tab:phase0-tests}
\resizebox{\textwidth}{!}{
\begin{tabular}{|p{0.6cm}|p{6.2cm}|p{9.0cm}|p{1.6cm}|}
\hline
\textbf{Grp} & \textbf{Test name} & \textbf{What it checks} & \textbf{Result} \\
\hline
A & \texttt{test\_one\_decision\_per\_zone\_per\_lap} & Each (driver, zone, lap) tuple appears at most once per run in the decision event log & Pass \\
\hline
A & \texttt{test\_cooldown\_enforcement} & Any two sequential attempt events for the same driver are separated by at least 5.0~s of simulated time & Pass \\
\hline
A & \texttt{test\_terminal\_reward\_correctness} & \texttt{outcome\_raw} equals starting position minus final position for every driver & Pass \\
\hline
A & \texttt{test\_reward\_sum\_identity} & \texttt{reward\_component\_sum\_error} is below $10^{-4}$ for every decision event & Pass \\
\hline
A & \texttt{test\_inactive\_components\_zero} & \texttt{pace} and \texttt{tyre\_pit} components are exactly 0.0 in every event at low complexity & Pass \\
\hline
B & \texttt{test\_aggression\_increases\_probability} & Mean success probability across 1000 samples satisfies AGGRESSIVE $>$ NORMAL $>$ CONSERVATIVE at fixed zone and gap & Pass \\
\hline
B & \texttt{test\_gap\_decreases\_probability} & Closer gap yields higher mean success probability at 1000 samples & Pass \\
\hline
B & \texttt{test\_zone\_difficulty\_scales\_probability} & Lower difficulty zone yields higher mean success probability at 1000 samples & Pass \\
\hline
C & \texttt{test\_jsonl\_schema} & Required top-level keys present in every race results JSONL record & Pass \\
\hline
C & \texttt{test\_decision\_event\_schema} & Required per-decision keys present including \texttt{reward\_component\_sum\_error} and \texttt{success\_probability} & Pass \\
\hline
C & \texttt{test\_component\_totals\_match\_events} & Per-driver \texttt{reward\_component\_totals} match the sum of individual decision events for decision-time components & Pass \\
\hline
\end{tabular}
}
\end{table}

\subsection{Defects found and corrected}

Two defects were identified during Phase~0 and corrected before any Phase~1 runs were executed.

The first defect was in the base agent heuristic. The \texttt{default\_policy} function in \texttt{src/base\_agents.py} reads the gap to the car ahead via \texttt{getattr(driver, "gap\_to\_ahead", None)}. The \texttt{DriverState} class did not initialise this attribute, so the value was always \texttt{None}. This collapsed the base agent's overtake attempt probability to approximately 19\% at La Source instead of the intended 77\%, making the base agent effectively passive. The defect was corrected by setting \texttt{driver.gap\_to\_ahead = gap\_distance} in the simulator's \texttt{\_check\_overtakes} method before the agent action call. After correction the base agent attempted overtakes at a frequency consistent with its heuristic policy parameters.

The second defect was a scoping error in \texttt{test\_one\_decision\_per\_zone\_per\_lap}. The deduplication set was initially keyed on driver name alone rather than on the (run number, driver name) pair, causing legitimate across-run repetitions to trigger false failures. The test was corrected to scope deduplication within each individual run.

\subsection{Phase 0 gate outcome}

All twelve tests passed. No \texttt{reward\_component\_sum\_error} exceeded $10^{-4}$. The stochastic calibration ordering assertions held across all 1000-sample replicates. The JSONL schema contained all required keys. Gate~G0 was cleared and Phase~1 proceeded.

% ---------------------------------------------------------------
\section{Phase 1: Single-Agent Smoke Evaluation}
% ---------------------------------------------------------------

\subsection{Objective and configuration}

Phase~1 confirmed that a learning signal existed before the full Phase~2 benchmark matrix was run. All five smoke runs used a vanilla DQN agent trained for 200 episodes against a single heuristic base agent at low complexity on the Spa circuit. The epsilon decay was set to 0.995, producing $\varepsilon \approx 0.37$ by episode 200. Each run was evaluated over 20 races per seed across three evaluation seeds (101, 202, 303). The three s0 runs used independent training seeds (101, 202, 303). The s1 and s2 runs used training seed 101 only.

The stochasticity levels used Gaussian noise on the overtake success probability. s0 used noise standard deviation 0.0 (deterministic). s1 used 0.02. s2 used 0.05.

\subsection{s0 results: three training seeds}

\begin{table}[htbp]
\centering
\small
\caption{Phase 1 s0 smoke results across three training seeds (60 eval races per seed set)}
\label{tab:phase1-s0}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
\textbf{Train seed} & \textbf{Win rate} & \textbf{CI95 low} & \textbf{CI95 high} & \textbf{Position delta} & \textbf{Seed var.} & \textbf{DNF rate} \\
\hline
101 & 0.783 & 0.678 & 0.888 & 0.567 & 0.0033 & 0.0 \\
202 & 0.767 & 0.659 & 0.875 & 0.533 & 0.0058 & 0.0 \\
303 & 0.783 & 0.678 & 0.888 & 0.567 & 0.0033 & 0.0 \\
\hline
\end{tabular}
\end{table}

All three seeds exceeded the Gate~G1 criterion of win rate above 0.50 with CI95 lower bound above 0.50. Starting position exposure was perfectly balanced at 30 races from each of positions 1 and 2 in every run. No DNFs occurred. Gate~G1 was cleared.

The three seeds produced closely grouped results, with seed variance between 0.003 and 0.006. This indicated convergence to a similar policy attractor across independent training runs.

\subsection{Risk behaviour at s0}

\begin{table}[htbp]
\centering
\small
\caption{Risk level counts and zone-level attempt rates at s0 (train seed 101, 60 eval races)}
\label{tab:phase1-s0-risk}
\begin{tabular}{|l|c|c|c|c|c|}
\hline
\textbf{Zone} & \textbf{Difficulty} & \textbf{Decisions} & \textbf{Attempt rate} & \textbf{Success rate} & \textbf{Avg reward} \\
\hline
La Source       & 0.2 & 30 & 1.00 & 0.867 & $+0.127$ \\
Raidillon       & 0.9 & 28 & 1.00 & 0.143 & $-0.021$ \\
Les Combes      & 0.7 & 46 & 0.870 & 0.300 & $+0.013$ \\
Bruxelles       & 0.7 & 13 & 1.00 & 0.231 & $+0.015$ \\
\hline
\end{tabular}
\end{table}

Risk level counts at s0 were CONSERVATIVE~60, NORMAL~42, AGGRESSIVE~10. The agent showed a strong preference for cheaper-failure strategies at s0. Only four of nine configured zones appeared in any decision event. Zones at greater track distances from the start were not reached in sufficient proximity situations to trigger decisions within the two-car five-lap race format.

The agent attempted at Raidillon (difficulty 0.9) 100\% of the time despite a 14\% success rate and a net-negative average reward. This indicated that the policy had not learned to discriminate HOLD decisions at high-difficulty zones within 200 training episodes.

\subsection{Stochasticity degradation: s0, s1, and s2}

\begin{table}[htbp]
\centering
\small
\caption{Phase 1 stochasticity comparison (train seed 101, 60 eval races per level)}
\label{tab:phase1-stoch}
\begin{tabular}{|l|c|c|c|c|c|c|}
\hline
\textbf{Level} & \textbf{Win rate} & \textbf{CI95 low} & \textbf{CI95 high} & \textbf{Tactical raw total} & \textbf{Outcome raw total} & \textbf{Seed var.} \\
\hline
s0 & 0.783 & 0.678 & 0.888 & $-11.0$ & $+17.0$ & 0.0033 \\
s1 & 0.783 & 0.678 & 0.888 & $-17.5$ & $+17.0$ & 0.0033 \\
s2 & 0.750 & 0.640 & 0.860 & $-27.5$ & $+15.0$ & 0.0100 \\
\hline
\end{tabular}
\end{table}

\begin{table}[htbp]
\centering
\small
\caption{Risk level counts across stochasticity levels (train seed 101, 60 eval races per level)}
\label{tab:phase1-risk-levels}
\begin{tabular}{|l|c|c|c|c|c|}
\hline
\textbf{Level} & \textbf{CONSERVATIVE} & \textbf{NORMAL} & \textbf{AGGRESSIVE} & \textbf{Decisions} & \textbf{Attempt rate/run} \\
\hline
s0 & 60 & 42 & 10 & 118 & 1.87 \\
s1 & 19 & 26 & 48 & 110 & 1.55 \\
s2 & 18 & 30 & 49 &  97 & 1.62 \\
\hline
\end{tabular}
\end{table}

% ---------------------------------------------------------------
\section{Phase 1 findings and implications}
% ---------------------------------------------------------------

\subsection{Win rate floor: La Source structural dominance}

The win rate was flat from s0 to s1 (0.783 in both cases). This was not a sign of robustness. The dominant overtaking zone at La Source (difficulty 0.2) produced an 87--90\% success rate in all three levels with 100\% attempt rate. Adding a noise standard deviation of 0.02 to a near-certain probability event is insufficient to flip race outcomes at La Source. The s1 noise level fell below the La Source erosion threshold, making the s0-to-s1 comparison uninformative about true policy robustness.

The tactically informative metric was the win rate from position-two starts. Stripping out the structural position-one advantage, the agent won approximately 57\% of P2-start races at both s0 and s1, falling to 50\% at s2. This 7 percentage point reduction at s2 was the genuine sensitivity signal.

\subsection{AGGRESSIVE flip as a DQN overestimation indicator}

The risk level distribution reversed completely between s0 and s1. At s0 the agent used CONSERVATIVE in 53\% of attempts. At s1 and s2 it used AGGRESSIVE in 52--51\% of attempts. This reversal persisted across stochasticity levels once noise was introduced.

The root cause was Q-value overestimation under high-variance rewards. Vanilla DQN uses a max operator in the Bellman update, which preferentially propagates optimistic samples. Under noisy outcomes, AGGRESSIVE occasionally produced high-reward events. These upward noise samples inflated Q-values for AGGRESSIVE in training, biasing the policy toward the highest-penalty failure strategy at evaluation. The consequence appeared directly in the tactical component, which worsened from $-11.0$ at s0 to $-27.5$ at s2, a 2.5$\times$ increase despite fewer total decisions.

This overestimation pattern under noise is the documented failure mode that motivated Double~DQN and is empirically visible here. It provides direct motivation for the Phase~2 algorithm comparison.

\subsection{Zone discrimination collapse at s2}

At s0 the agent applied a partial HOLD policy at Les Combes (87\% attempt rate). At s1 this discrimination strengthened to 73\% attempt rate, and the Raidillon HOLD rate improved to 90\%. At s2 both zones reverted to 100\% attempt rate. The learned discrimination did not persist beyond moderate noise. This indicated that the Q-value signal was too weak at 200 training episodes to maintain zone-specific HOLD behaviour when outcome variance exceeded the signal from the penalty gradient.

\subsection{Seed variance as a robustness indicator}

Seed variance was stable at 0.0033 across s0 and s1, then tripled to 0.010 at s2. This was the first measurable instability signal and the entry point for the RQ1 robustness analysis. Policy sensitivity to training seed increased at s2, confirming that higher noise was beginning to disrupt convergence reliability even within the 200-episode budget.

\subsection{Phase 1 gate summary}

\begin{table}[htbp]
\centering
\small
\caption{Phase 1 gate G1 checklist}
\label{tab:phase1-gate}
\begin{tabular}{|p{8.5cm}|p{2.0cm}|}
\hline
\textbf{Criterion} & \textbf{Outcome} \\
\hline
At least one of three s0 seeds shows win rate above 0.50 & Met (all three) \\
All CI95 lower bounds above 0.50 at s0 & Met (lowest: 0.659) \\
Fairness violations equal zero (start position imbalance) & Met (ratio 0.0) \\
s1 and s2 show interpretable degradation direction & Met (see note) \\
\hline
\end{tabular}
\end{table}

The degradation direction was interpretable but structurally attenuated. The headline win rate was flat at s1 due to La Source dominance, and only the s2 comparison showed a reduction. The behavioural degradation signal (AGGRESSIVE flip, zone discrimination collapse, rising tactical penalty, increased seed variance) was present and mechanistically consistent across s1 and s2. Gate~G1 was cleared and Phase~2 proceeded with this understanding recorded.

\end{appendices}
