This project investigates how incentive structure shapes emergent strategy when multiple reinforcement learning agents train together in a stochastic environment. The domain is Formula 1 overtaking, modelled as discrete decisions at zones along the Spa-Francorchamps circuit. A reward-sharing coefficient controls how much each agent values a teammate's performance alongside its own, ranging from purely self-interested to fully cooperative. By keeping other variables constant and varying this coefficient, incentive design is isolated from algorithm choice, game structure, and team size.

Four variants of the Deep Q-Network family are compared as the learning substrate. Under concurrent competitive training, algorithm choice determines whether stable strategic behaviour emerges at all. Weaker variants collapse into passivity in close to half of all trials. The strongest variant, combining prioritised experience replay with double-target correction, avoids collapse entirely and develops clean complementary roles resembling the hawk-dove equilibrium from evolutionary game theory, with one agent taking aggressive zones and the other conservative ones without any shared coordination mechanism.

Introducing cooperative reward into a purely competitive game produces a sharp breakdown. Below a threshold, partial alignment does not destabilise learning. Above it, agents converge to passivity in the majority of trials because successful overtakes turn into net penalties. Changing the game structure so both agents can improve simultaneously reverses the picture. With a shared adversary to outperform together, the same cooperative incentive produces an 83 per cent improvement in joint performance, driven by lifting the weaker agent rather than uniformly raising both.

Scaling to more agents eliminates this advantage. The cooperative gain that emerges at three agents vanishes abruptly at four, and no coefficient value recovers it. Curriculum training, which establishes individual policies before introducing cooperative reward, restores much of the lost performance yet never exceeds the competitive baseline. Two distinct failure modes become separable. Training-order collapse is resolved by scheduling. Credit-assignment failure at scale is not, and points to architectural change beyond the independent-learning framework.

A zero-shot large language model given the race state as a natural-language description matches the best-trained agent across all tested conditions, through qualitatively different strategies. This suggests the environment admits a short near-optimal policy that domain reasoning alone can capture. An alternative reward formulation designed to credit each agent only for its own counterfactual contribution fails both analytically and empirically. The mean-field approximation used inverts the cooperative gradient, penalising teammate improvement rather than rewarding own actions.

The central finding is that incentive design and agent count jointly determine whether cooperation produces advantage or harm. In small teams, careful alignment of incentives produces real cooperative gains. In larger teams, reward-level interventions alone are insufficient.