\chapter{Introduction and Background Research}
\label{chapter1}

\section{Introduction}

Formula 1 is one of the most strategically complex sports in the world. Beneath the speed lies a continuous decision problem. At every corner, drivers and engineers weigh whether to attack, defend, or wait. A single choice made a fraction of a second too early or too late can determine the outcome of the race.

What makes this genuinely difficult is that no decision exists in isolation. When two cars fight for position at the same braking zone, the outcome depends on the choices of both drivers, not just one. Race strategy is, at its core, a problem of multiple decision-makers interacting in a shared environment under uncertainty.

This dissertation investigates that problem from a specific angle. Rather than asking whether an artificial agent can win races, it asks how the rules of what agents are rewarded for shape the strategies they develop. In particular, it examines how changing the degree to which two agents share each other's reward alters the kind of strategic behaviour that emerges when they learn simultaneously.

This chapter builds the conceptual foundation for this investigation in four steps. First, it introduces the field of competitive autonomous decision-making and the algorithms that power it. Second, it surveys what existing research has established within this field. Third, it identifies the niche that existing work has not yet explored, specifically the role of incentive structure in shaping emergent social strategy between concurrent learners. Fourth, it states the research gap and the specific contribution of this project. A reader new to reinforcement learning or game theory should be able to follow the reasoning to the research questions by the end of this chapter.

\section{Learning to compete: reinforcement learning in complex environments}

Reinforcement learning is a style of machine learning in which an agent learns by doing. Rather than memorising a set of correct answers, the agent interacts with an environment, takes actions, and receives a numerical signal called a reward that tells it how well it is doing. Over many repeated interactions, the agent adjusts its behaviour to collect more reward \citep{ribeiro2002rlagents}.

This approach suits race strategy for a direct reason. There is no fixed correct overtaking policy that a textbook could supply. The right choice at any given moment depends on the track location, the gap to the car ahead, the relative speeds of both cars, and how conditions are likely to change over the remaining laps. These factors combine in too many ways for a hand-coded rule set to cover. Reinforcement learning allows an agent to experience many races, learn what worked and what did not, and build a policy from accumulated experience.

The specific form used in this project is value-based learning. In this approach, the agent learns to estimate the long-run expected reward from each possible action in each possible situation. This estimate is called the Q-value. An agent with accurate Q-values can select the action that looks most promising in the long run, not just the one that produces the largest immediate return \citep{mnih2015human}. In a race, this distinction matters. An aggressive overtake attempt might succeed immediately, but if it risks a collision or depletes tyre life, it can be a poor decision for the race as a whole. A well-trained Q-value function captures these trade-offs.

The DQN family implements value-based learning using deep neural networks to approximate Q-values from raw state observations. The original DQN established that this combination can achieve strong performance across a range of complex environments \citep{mnih2015human}. Later variants each addressed a specific weakness. Double DQN corrects a systematic tendency to overestimate Q-values when outcomes are noisy \citep{vanhasselt2016double}. Dueling DQN separates the estimate of how good a state is from the estimate of which action is best in that state, helping in situations where most actions have similar consequences \citep{wang2016dueling}. Prioritised experience replay focuses learning on the most informative past experiences rather than sampling them uniformly \citep{schaul2016prioritized}. Rainbow combines several of these improvements into a unified framework \citep{hessel2018rainbow}. These four variants form the algorithmic core of this investigation. The reasoning behind this choice is developed further in Section~\ref{sec:dqn-foundations}.

Single-agent reinforcement learning assumes the environment follows stable rules. The agent is the only thing that changes. If a policy works well in one episode, it can reasonably be expected to work under the same conditions in the next.

Multi-agent reinforcement learning breaks this assumption. When several agents learn at the same time, each agent's environment is shaped by the policies of all the others. As those policies change, the effective environment experienced by any one agent changes too. An action that produced a high reward yesterday may produce a very different result today, because the opponent has since learned to respond differently. This property is called non-stationarity, and it is the defining technical challenge of multi-agent reinforcement learning \citep{busoniu2008survey,hernandezleal2019survey}.

Non-stationarity creates a specific problem for Q-value estimation. Q-values are supposed to reflect the long-run expected reward from an action. But if opponents keep changing their behaviour, the true expected reward from any action keeps changing too. Q-values trained against yesterday's opponent may be badly wrong about today's. This can cause learning to destabilise, with policies cycling or collapsing rather than converging \citep{foerster2017stabilising,matignon2012independent}.

In racing, non-stationarity is not a theoretical complication. It is part of what makes strategy interesting. A driver who always attacks at the same corner becomes predictable. A driver who varies their approach keeps the opponent uncertain. When both drivers are learning simultaneously, the interesting question is whether their policies settle into stable complementary strategies, or whether mutual adaptation produces instability and collapse. This question sits at the centre of the present investigation.

\section{What the research has established}
\label{sec:litreview}

Research in this space spans three intersecting bodies of work. The first covers motorsport simulation and planning. The second covers reinforcement learning in competitive settings, including the reproducibility conditions required for trustworthy comparisons. The third covers multi-agent reinforcement learning, where the specific challenges of non-stationarity and coordination have been studied in depth. A structured comparison of key related studies is provided in Appendix Table~\ref{tab:related-work-comparison}.

\subsection{Motorsport simulation and planning}

Early motorsport strategy research focused on fast simulation for decision support. \citet{heilmeier2018racesim} built a lap-wise race simulation modelling tyre degradation, fuel effects, pit stops, and overtaking processes using publicly accessible race data. The strength of this work is engineering pragmatism. The simulation is grounded in real-world observations, giving the outputs practical credibility. The limitation for the present project is that the simulator was designed to evaluate candidate strategies rather than compare how different learning algorithms develop distinct behaviours under matched conditions.

\citet{heilmeier2020virtual} extended this line of work with neural network approximations, showing that data-driven surrogates can accelerate strategy exploration. \citet{piccinotti2021online} investigated online planning for Formula 1, demonstrating that planning methods can produce competitive decisions in dynamic race conditions. Planning-focused studies typically evaluate decision quality against a fixed simulator rather than comparing the emergent strategies of multiple concurrently learning agents.

Formula E research adds a parallel thread. \citet{liu2020formulae} combined neural modelling with Monte Carlo tree search for energy strategy, showing how sequential decision-making under resource constraints can be framed as a multi-stage problem. More recent multi-car Formula E work demonstrates the relevance of reinforcement learning in genuinely competitive settings \citep{liu2024formulae}. The most recent motorsport multi-agent work, \citet{fieni2026learning}, reports learning-based multi-agent race strategies in Formula 1 contexts, signalling growing interest in this direction.

Taken together, this body of work shows that abstracted simulators can produce meaningful tactical insight without full physical fidelity. Reinforcement learning is an accepted framework for racing strategy. What no existing motorsport study has done is systematically vary the incentive structure between concurrent learning agents, or test how reward alignment between agents alters the character of the strategies that emerge.

\subsection{Reinforcement learning in competitive environments}

Before examining multi-agent dynamics, it is important to understand what reinforcement learning research has established about learning under uncertainty.

\citet{ribeiro2002rlagents} provides a foundational framing. Reinforcement learning is learning through interaction with an uncertain environment. This applies directly to race strategy, where policies must handle both environmental uncertainty from stochastic overtake outcomes, and strategic uncertainty from an opponent whose behaviour is not known in advance.

A useful illustration from an adjacent domain appears in \citet{groeneveld2023recommerce}, which shows that self-learning agents in competitive market environments must develop strategies that are robust to opponent adaptation, not just to fixed opponent behaviour. The methodological lesson transfers directly. Benchmarks that evaluate agents only under fixed, predictable conditions can produce results that do not survive contact with a genuine adapting opponent. This motivates the stochasticity sweep used in the present investigation.

Reproducibility is a further critical constraint. \citet{henderson2018deep} showed that random seeds, implementation choices, and reporting conventions can change the apparent conclusions of reinforcement learning experiments. Two studies using the same algorithm under different evaluation procedures can produce contradictory rankings. \citet{agarwal2021statistical} extended this finding, showing that mean-only reporting is unreliable for ranking algorithms and that confidence intervals and across-seed variance are necessary for trustworthy comparison. These findings shape the evaluation design of this project. Matched compute budgets, fixed seeds, confidence interval reporting, and behavioural diagnostics are treated as mandatory controls rather than optional additions. \citet{sankary2018stochastic} demonstrated that uncertainty-aware evaluation procedures reduce overfitting to narrow scenario sets and produce more generalisable conclusions. Testing across multiple stochasticity levels, as this project does, follows the same principle.

\subsection{Multi-agent reinforcement learning}

\citet{busoniu2008survey} provides the foundational framework for multi-agent reinforcement learning. The approach differs from single-agent learning along three dimensions. Outcomes depend on the joint actions of all agents, not just one. Agents typically observe only their own local state, not the complete global picture. And because all agents are learning simultaneously, the environment experienced by any one agent changes over time. Applying single-agent algorithms in multi-agent settings without accounting for these properties leads to unstable or degenerate outcomes.

\citet{hernandezleal2019survey} surveys the deep learning era and identifies a central tension. When agents are purely competitive, each maximising its own reward at the expense of others, game-theoretic equilibrium analysis is the natural framework. When agents are purely cooperative, maximising a shared objective, coordination becomes the primary challenge. When agents are mixed, partly competing for individual rewards and partly sharing consequences of each other's outcomes, the landscape is richer and less well understood. This mixed setting is precisely the one examined in the present project.

\citet{tan1993multiagent} established that independent learners in cooperative games can outperform a single centralised agent when the environment is sufficiently complex, but that performance degrades as the number of agents grows. \citet{panait2005cooperative} survey cooperative multi-agent reinforcement learning and identify credit assignment as the central unsolved problem in team reward settings. \citet{wang2022irat} propose mixing individual and team rewards to balance credit assignment against cooperative alignment, which is the mechanism examined through the alpha coefficient in this project.

\citet{foerster2017stabilising} identifies a specific failure mode when independent Q-learners operate in the same environment. Each agent's replay buffer fills with experiences generated under the opponent's past policy. When that policy changes, the stored experiences become outdated. Q-values trained on stale experience can misdirect learning and produce instability. \citet{matignon2012independent} surveys this class of failure more broadly, documenting pathological convergence patterns across a range of multi-agent settings. In the present project, this failure mode produces what is termed degenerate collapse. The trailing agent's replay buffer becomes dominated by unsuccessful attempts, Q(HOLD) climbs above Q(ATTEMPT) across all zones, and the agent converges to complete passivity. This pattern is analysed in detail in Chapter~\ref{chapter4}.

Centralised training with decentralised execution represents the natural next step beyond independent learners. In this framework, agents are trained using global state information that is unavailable at execution time, allowing them to learn coordinated policies that are still executed from local observations alone. \citet{foerster2018coma} demonstrated this approach using a counterfactual advantage function that credits individual agent contributions to joint outcomes. Value factorisation methods including VDN \citep{sunehag2018vdn} and QMIX \citep{rashid2018qmix} decompose a joint Q-value into per-agent contributions, preserving decentralised execution whilst enabling cooperative optimisation. These methods represent a natural progression beyond the independent learner baseline used in this dissertation and are positioned as future work in Chapter~\ref{chapter4}.

The mixed cooperative-competitive setting is addressed directly by \citet{lowe2017maddpg}, which introduced MADDPG for environments where some agents cooperate and others compete. MADDPG uses centralised critics trained with knowledge of all agents' actions to stabilise learning in these complex settings. Whilst MADDPG is not implemented in this dissertation, it provides evidence that mixed incentive environments require specialised training structures, and that training architecture interacts with incentive design in ways that purely competitive or purely cooperative settings alone do not reveal.

\section{The niche: social strategy and incentive structure}
\label{sec:social-dilemmas}

The reviewed work establishes that multi-agent reinforcement learning is technically mature and that motorsport simulation is a credible research domain. But a specific angle within this space has received almost no attention. That angle is the role of incentive structure in determining what social strategies emerge between concurrent learners.

\subsection{Social dilemmas and the emergence of cooperation}

A social dilemma is a situation in which the choice that is best for an individual agent leads to an outcome that is worse for everyone, compared to what cooperation would produce \citep{leibo2017multiagent}. The prisoner's dilemma is the classic illustration. Two suspects are questioned separately. If both stay silent, both receive a light sentence. If one confesses while the other stays silent, the confessor goes free and the other receives a heavy sentence. If both confess, both receive a moderate sentence. The individually rational choice is always to confess, regardless of what the other person does. But when both agents follow this logic, both end up worse off than if they had stayed silent. Rational individual choices produce a collectively harmful outcome.

\citet{leibo2017multiagent} formalised this concept within multi-agent reinforcement learning. They showed that sequential social dilemmas arise naturally when agents interact repeatedly in shared environments with overlapping resource interests. Independent learners frequently converge to defection even when cooperation would yield higher collective reward.

Social dilemmas appear naturally in team racing. Consider two cars from the same team. Both drivers have individual incentives to finish as high as possible. But the team's overall result depends on both cars finishing well. If one car sits behind the other despite being faster, the team might benefit from the leading car ceding position. The leading car's individual incentive is to hold position. The tension between team-level welfare and individual reward is a genuine social dilemma. Formula 1 teams navigate this through team orders, driver contracts, and explicit agreements.

In a simulated setting, an equivalent dilemma emerges when two agents both target the easiest overtaking zone on every lap. They collectively reduce each other's opportunity by competing for the same resource. A secondary zone could be accessed with high collective value if each agent committed to a different territory. This mirrors the tragedy of the commons, where shared resources are overexploited by individually rational agents at the cost of collective efficiency \citep{nowak2006five}.

\citet{hughes2018inequity} showed that modifying agents' reward functions to incorporate partial concern for teammates' outcomes substantially increases cooperative behaviour in sequential dilemmas. When each agent receives a portion of its teammate's reward, competing aggressively for the teammate's territory becomes costly to the self. Conditions for complementary specialisation can then emerge naturally. One agent commits to a primary zone, the other develops a distinct territory, and both benefit from reduced direct competition.

\citet{nowak2006five} identifies five mechanisms through which cooperation is sustained in populations of interacting agents. These are kin selection, direct reciprocity, indirect reciprocity, network reciprocity, and group selection. In the present setting, the most relevant mechanism is direct reciprocity. Agents that repeatedly interact with the same opponent can sustain cooperation through repeated best-response dynamics. Partial incentive alignment strengthens this mechanism by making cooperation structurally attractive even when reciprocity signals are noisy.

\subsection{The reward-sharing coefficient as a research variable}

This project introduces a reward-sharing coefficient, denoted alpha, as the central experimental variable. Alpha controls how much each agent values its teammate's outcome alongside its own. When alpha is zero, agents are purely competitive. Each cares only about its own finishing position. When alpha is one, agents are fully cooperative. Each cares equally about both outcomes. At values in between, agents are partially aligned. Individual incentive is moderated by a degree of concern for the teammate's result.

The prediction from the social dilemma literature is that changing alpha changes not just win rates, but the qualitative character of the strategies that emerge \citep{hughes2018inequity,nowak2006five}. It should affect how agents partition territory across overtaking zones, how their risk profiles diverge from one another, and how stable those patterns remain as the race environment becomes more uncertain.

\subsection{Why the DQN family is the right substrate}
\label{sec:dqn-foundations}

The choice of the DQN family as the algorithmic substrate for the incentive-regime comparison was made on five specific grounds.

The first is action-space fit. The decision space in this simulator is discrete. At each overtaking zone, an agent chooses to hold position or attempt an overtake at one of three risk levels. This compact discrete action space suits Q-value methods, which maintain separate value estimates for each action \citep{zhu2021actionspaces,mnih2015human}.

The second is off-policy replay efficiency. Value-based methods with experience replay can reuse past transitions to learn from rare but informative events more efficiently than methods that discard experience after a single update. In a five-lap race with sparse decision points, this sample efficiency matters \citep{mnih2015human,schaul2016prioritized}.

The third is Q-value overestimation under noise. In stochastic environments, vanilla Q-learning systematically overestimates action values because the maximum operation in the Bellman target is biased upward under variance. Double DQN corrects this by separating action selection from value evaluation in the target computation \citep{vanhasselt2016double}. This is particularly relevant because noisy overtake outcomes inflate the apparent Q-value of aggressive actions, producing the AGGRESSIVE flip documented in Chapter~\ref{chapter4}.

The fourth is value decomposition relevance. Dueling architecture separates state value from action advantage \citep{wang2016dueling}. This benefits learning in states where the choice of action matters little, such as at a very difficult zone where holding position is almost always correct regardless of risk level chosen.

The fifth is within-family comparability. By comparing variants within the same algorithmic family under identical training infrastructure, the effect of each specific mechanism can be isolated cleanly. This controlled ablation is essential for the incentive-regime analysis. It ensures that differences in emergent strategy across alpha values can be attributed to incentive structure rather than to algorithm-level instability \citep{henderson2018deep,agarwal2021statistical}.

DQN-family independent learners constitute the standard first baseline in the multi-agent reinforcement learning literature for value-based settings \citep{hernandezleal2019survey}. Establishing which variant is resistant to collapse in the competitive baseline condition is a prerequisite for the incentive-regime comparison. Without a stable algorithmic substrate, differences in emergent strategy across alpha values cannot be attributed to incentive structure rather than to algorithm-level fragility.

\section{The research gap and this project's contribution}

A clear gap exists at the intersection of the four bodies of work reviewed above. No existing motorsport study has systematically varied the incentive structure between concurrently learning agents in a stochastic simulator. No existing MARL study has examined how varying the degree of shared incentive between concurrent learners in a competitive environment alters the character of the strategies that emerge, or whether such variation produces genuine cooperative advantage as distinct from redistributed competitive outcomes. No existing study combines these questions with a matched-budget evaluation protocol that enforces confidence interval reporting and behavioural diagnostics alongside raw win rate.

This project addresses that gap through six linked contributions. First, it frames race strategy as a social dilemma in which incentive structure determines emergent behaviour, not algorithm choice alone. Second, it implements a reward sharing coefficient spanning the full spectrum from purely competitive to fully cooperative and measures its effect in both zero-sum and non-zero-sum game structures. Third, it identifies a sharp scaling boundary at which cooperative incentive transitions from beneficial to harmful, and characterises this boundary as structural rather than parametric through a controlled ablation across agent counts. Fourth, it separates two previously conflated failure modes through curriculum alpha scheduling. One failure mode is a training-order problem that curriculum resolves. The other is a credit assignment problem that requires architectural solutions. Fifth, it provides a mathematical proof that a mean-field approximation of difference rewards \citep{wolpert2002optimal} inverts the cooperative gradient, producing competitive rather than cooperative incentives. Sixth, it establishes a zero-shot large language model as a non-learning baseline that matches the best trained agent within two per cent win rate, raising questions about what value-based learning adds in low-complexity discrete racing environments.

The testbed is an F1-inspired simulator using the Spa-Francorchamps circuit, with nine overtaking zones derived from real historical telemetry, a stochastic overtake resolution mechanism, and a configurable reward contract. This setting captures the essential structure of competitive multi-agent decision-making under uncertainty whilst remaining controlled enough for reproducible scientific comparison. A faithful simulation of tyre degradation, aerodynamic wake effects, and pit-crew timing would obscure the behavioural dynamics of interest rather than illuminate them. The two-agent five-lap discretised race provides the appropriate scale for establishing clean baseline results before more complex dynamics are introduced.

Three levels of environmental uncertainty are tested. Level s0 produces fully deterministic outcomes given the same state and action. Level s1 adds small Gaussian noise to overtake probabilities, introducing modest outcome variance. Level s2 adds moderate noise sufficient to flip borderline outcomes and meaningfully stress learned policies. Testing across all three levels determines whether findings about incentive structure hold as the environment becomes more unpredictable.

The novelty of this contribution is both methodological and substantive. Methodologically, the project provides a reproducible incentive regime comparison framework adaptable to other competitive simulation domains. Substantively, it provides controlled empirical evidence of how reward sharing between concurrent learners shapes the transition between competitive and cooperative strategy, and identifies the conditions under which that transition produces genuine joint benefit rather than pathological convergence.

\section{Aim, objectives, and research questions}

\subsection{Aim}

This project investigates how the degree of shared incentive between concurrent reinforcement learning agents shapes the strategies they develop. Two further questions follow naturally. Do those strategies remain stable as environmental uncertainty increases? Does cooperative incentive produce genuine coordination, or does it cause learning to collapse? A stochastic racing simulator built for this purpose provides the controlled setting in which incentive structure is the sole experimental variable.

\subsection{Objectives}

\begin{enumerate}
    \item Design and validate a stochastic racing simulator suitable for controlled multi-agent reinforcement learning experiments, with configurable complexity, reproducible seeding, and structured behavioural telemetry.
    \item Evaluate DQN-family algorithms in single-agent and multi-agent settings to identify which variant provides a stable learning substrate for the incentive regime comparison.
    \item Characterise the emergent strategic behaviour of two competing agents under purely competitive conditions, establishing a behavioural baseline against which cooperative incentive effects can be measured.
    \item Systematically vary the reward sharing coefficient between agents and measure its effect on strategic differentiation, competitive balance, and learning stability across multiple levels of environmental uncertainty.
    \item Assess whether shared incentive produces genuine cooperative advantage in the form of improved joint performance against a common adversary, or whether it merely redistributes outcomes within a fixed interaction.
    \item Characterise the agent-count boundary at which cooperative incentive transitions from beneficial to harmful, and determine whether this boundary is parametric or structural.
    \item Evaluate difference rewards as an alternative cooperative formulation and establish whether a mean-field approximation preserves the cooperative gradient property of the original \citet{wolpert2002optimal} formulation.
\end{enumerate}

\subsection{Research questions and hypotheses}

\textbf{RQ1.} What patterns of strategic behaviour emerge when independent reinforcement learning agents train concurrently in a shared stochastic environment, and how sensitive are those patterns to environmental noise?

\textbf{H1.} Concurrent training produces interpretable strategic structure including asymmetric risk preferences and differentiated zone usage. These patterns arise from Q-value divergence between agents exposed to different positional histories during training. Higher stochasticity flattens Q-value gradients across zones by reducing the reliability of any single zone's expected return, which weakens the basis for stable specialisation. Phase~3 confirms this. Rainbow-lite trials show hawk-dove equilibria and zone specialisation in all nine trials, while vanilla DQN produces degenerate collapse in four of nine.

\textbf{RQ2.} How does systematically varying the reward sharing coefficient between agents alter the emergent strategy profile, and is there a threshold beyond which shared incentive destabilises rather than improves learning?

\textbf{H2.} Increasing the reward sharing coefficient from zero initially increases strategic differentiation, as agents with partially aligned incentives face mild pressure to avoid direct competition for the same resources. Beyond a critical threshold, the shared component of the reward signal dominates the individual component, inverting the incentive to improve own position. This produces convergence to passive equilibria rather than cooperation. The transition is expected to be sharp rather than gradual. Phase~4 confirms the sharp transition. Collapse rates jump from 11 per cent at alpha of 0.5 to 89 per cent at alpha of 0.75 in zero-sum games.

\textbf{RQ3.} Does shared incentive produce genuine cooperative advantage, measured as improved joint performance against a common adversary, or does it merely redistribute outcomes within a fixed interaction?

\textbf{H3.} In a zero-sum interaction between two agents, shared incentive cannot produce genuine cooperative advantage because the sum of outcomes is structurally fixed. Introducing a common adversary breaks this constraint. Under non-zero-sum conditions, partial incentive alignment is predicted to increase the rate at which both learning agents outperform the adversary. Zone differentiation reduces wasted competitive effort and allows each agent to exploit its strongest territory. Fully cooperative incentive may restore pathological convergence even in the non-zero-sum setting if the shared signal still dominates individual learning gradients. Phase~5 confirms genuine cooperative advantage at alpha of 0.75 in the non-zero-sum setting, with an ablation proving game structure rather than evaluation methodology as the causal variable.

The next chapter describes the simulator, the complexity tiers, the stochasticity pathway, the reward contract, and the experimental protocol through which these questions are examined.
